### 2018-12-12 reports/thesis/chapters/5_inference.tex
Completed un/contaminated writing

\subsection{Uncontaminated Experiment}
The empirical null filter was demonstrated on an image where each pixel has i.i.d.~$\normal(0,1)$ test statistics. The test statistics after filtering should preserve its original distribution, or at least close to it. An experiment was conducted to investigate the empirical properties when filtering such as an image.

An example of a  $\addNumber{../figures/inference/noContamination/imageSize.txt}\times \addNumber{../figures/inference/noContamination/imageSize.txt}$ Gaussian image before and after filtering, with a kernel radius of \addNumber{../figures/inference/noContamination/radius.txt} is shown in Figure \ref{fig:inference_noContamination_images}. A quick glance shows little difference, in particular the QQ plot in Figure \ref{fig:inference_noContamination_qq} shows that the post filter test statistics looked reasonably Normal. Because the empirical null mean and empirical null standard deviation are non-Normal random variables, the post filtered test statistics will never be Normal and any Normality tests would be too strict for the purpose of this experiment.

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.49\textwidth]{../figures/inference/noContamination/allNullGaussianqq.eps}
	\caption{QQ plot of the test statistics of a filtered $\addNumber{../figures/inference/noContamination/imageSize.txt} \times \addNumber{../figures/inference/noContamination/imageSize.txt}$ Gaussian image after filtering with kernel radius \addNumber{../figures/inference/noContamination/radius.txt}.}
	\label{fig:inference_noContamination_qq}
\end{figure}

\begin{figure}[p]
	\centering
	\centerline{
	\begin{subfigure}[b]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../figures/inference/noContamination/allNullGaussianunfiltered.eps}
		\caption{Unfiltered}
	\end{subfigure}
	\begin{subfigure}[b]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../figures/inference/noContamination/allNullGaussianfiltered.eps}
		\caption{Filtered}
	\end{subfigure}
	}
	\centerline{
	\begin{subfigure}[b]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../figures/inference/noContamination/allNullGaussiannullmean.eps}
		\caption{Empirical null mean}
	\end{subfigure}
	\begin{subfigure}[b]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../figures/inference/noContamination/allNullGaussiannullstd.eps}
		\caption{Empirical null std}
	\end{subfigure}
	}
	\caption{A $\addNumber{../figures/inference/noContamination/imageSize.txt} \times \addNumber{../figures/inference/noContamination/imageSize.txt}$ Gaussian image before and after filtering with kernel radius \addNumber{../figures/inference/noContamination/radius.txt}. Also shown are the empirical null mean and empirical null std.}
	\label{fig:inference_noContamination_images}
\end{figure}

@@

\begin{figure}
	\centering
	\begin{subfigure}[b]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../figures/inference/allnull/Experiment_AllNullGaussianmean.eps}
		\caption{Post filter sample mean}
	\end{subfigure}
	\begin{subfigure}[b]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../figures/inference/allnull/Experiment_AllNullGaussianvariance.eps}
		\caption{Post filter sample variance}
	\end{subfigure}
	\caption{The sample mean and sample variance of the test statistics in a filtered Gaussian image of size $\addNumber{../figures/inference/allnull/Experiment_AllNullGaussianwidth.txt} \times \addNumber{../figures/inference/allnull/Experiment_AllNullGaussianheight.txt}$. The box plots summarise the \addNumber{../figures/inference/allnull/Experiment_AllNullGaussiannrepeat.txt} repeats of the experiment. The dotted lines show the critical region at the 2$\sigma$ level using standard tests for the sample mean and sample variance, assuming independence.}
	\label{fig:inference_allnullgaussian}
\end{figure}

The results are shown in Figure \ref{fig:inference_allnullgaussian}.

\afterpage{\clearpage}
\subsection{Contaminated Experiment}
Consider a $\addNumber{../figures/inference/contamination/imageSize.txt}\times \addNumber{../figures/inference/contamination/imageSize.txt}$ Gaussian image with some contamination. For example let the pixels be $Z_{x,y}\sim\normal(\mu_{0,x,y},\sigma_0^2)$ where the null mean is a linear function
\begin{equation}
\mu_{0,x,y} = \addNumber{../figures/inference/contamination/gradx.txt} (x-x_0) + \addNumber{../figures/inference/contamination/grady.txt} (y-y_0)
\end{equation}
where $(x_0,y_0)$ is the centre of the image and $\sigma_0=\addNumber{../figures/inference/contamination/nullstd.txt}$. There are no defects or non-null pixels defined here, therefore any positive detections are false.

@@

The QQ plot of the filtered test statistics, as shown in Figure \ref{fig:inference_contamination_qq}, shows that there may be a variance mismatch. A mismatch in the variance will increase or decrease the sensitivity of the test, a larger variance would make more test statistics more likely to be tested as positive.

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.49\textwidth]{../figures/inference/contamination/allNullPlaneqq.eps}
	\caption{QQ plot of the test statistics of a filtered $\addNumber{../figures/inference/noContamination/imageSize.txt} \times \addNumber{../figures/inference/noContamination/imageSize.txt}$ contaminated Gaussian image after filtering with kernel radius \addNumber{../figures/inference/noContamination/radius.txt}.}
	\label{fig:inference_contamination_qq}
\end{figure}

\begin{figure}[p]
	\centering
		\centerline{
			\begin{subfigure}[b]{0.49\textwidth}
				\includegraphics[width=\textwidth]{../figures/inference/contamination/allNullPlaneunfiltered.eps}
				\caption{Unfiltered}
			\end{subfigure}
			\begin{subfigure}[b]{0.49\textwidth}
				\includegraphics[width=\textwidth]{../figures/inference/contamination/allNullPlanefiltered.eps}
				\caption{Filtered}
			\end{subfigure}
		}
		\centerline{
			\begin{subfigure}{0.49\textwidth}
				\includegraphics[width=\textwidth]{../figures/inference/contamination/allNullPlanenullmean.eps}
				\caption{Empirical null mean}
			\end{subfigure}
			\begin{subfigure}{0.49\textwidth}
				\includegraphics[width=\textwidth]{../figures/inference/contamination/allNullPlanenullstd.eps}
				\caption{Empirical null std}
			\end{subfigure}
		}
	\caption{A $\addNumber{../figures/inference/contamination/imageSize.txt} \times \addNumber{../figures/inference/contamination/imageSize.txt}$ contaminated Gaussian image before and after filtering with kernel radius \addNumber{../figures/inference/contamination/radius.txt}. The contamination is such that the null distribution is $Z_{x,y}\sim\normal(\mu_{0,x,y},\sigma_0^2)$ where $\mu_{0,x,y} = \addNumber{../figures/inference/contamination/gradx.txt} (x-x_0) + \addNumber{../figures/inference/contamination/grady.txt} (y-y_0)$ and $\sigma_0=\addNumber{../figures/inference/contamination/nullstd.txt}$. Highlighted in red are significant pixels when controlling the FDR at the $2\sigma$ level. Also shown are the empirical null mean and empirical null std.}
	\label{fig:inference_contamination_images}
\end{figure}

@@

The results are shown in Figure \ref{fig:inference_allnullplane}.

\begin{figure}
	\centering
	\begin{subfigure}[b]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../figures/inference/allnull/Experiment_AllNullPlanemean.eps}
		\caption{Post filter sample mean}
	\end{subfigure}
	\begin{subfigure}[b]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../figures/inference/allnull/Experiment_AllNullPlanevariance.eps}
		\caption{Post filter sample variance}
	\end{subfigure}
	\caption{The sample mean and sample variance of the test statistics in a filtered contaminated Gaussian image of size $\addNumber{../figures/inference/allnull/Experiment_AllNullGaussianwidth.txt} \times \addNumber{../figures/inference/allnull/Experiment_AllNullGaussianheight.txt}$. The contamination is such that the null distribution is $Z_{x,y}\sim\normal(\mu_{0,x,y},\sigma_0^2)$ where $\mu_{0,x,y} = \addNumber{../figures/inference/contamination/gradx.txt} (x-x_0) + \addNumber{../figures/inference/contamination/grady.txt} (y-y_0)$ and $\sigma_0=\addNumber{../figures/inference/contamination/nullstd.txt}$. The box plots summarise the \addNumber{../figures/inference/allnull/Experiment_AllNullPlanenrepeat.txt} repeats of the experiment. The dotted lines show the critical region at the 2$\sigma$ level using standard tests for the sample mean and sample variance, assuming independence.}
	\label{fig:inference_allnullplane}
\end{figure}

>>>>

\subsection{Uncontaminated Experiment}

The empirical null filter was demonstrated on an image where each pixel has i.i.d.~$\normal(0,1)$ test statistics. The test statistics after filtering should preserve its original distribution, or at least close to it. An experiment was conducted to investigate the empirical properties when filtering such as an image.

An example of a  $\addNumber{../figures/inference/noContamination/imageSize.txt}\times \addNumber{../figures/inference/noContamination/imageSize.txt}$ Gaussian image before and after filtering, with a kernel radius of \addNumber{../figures/inference/noContamination/radius.txt}, is shown in Figure \ref{fig:inference_noContamination_images}. A quick glance shows little difference, in particular the QQ plot in Figure \ref{fig:inference_noContamination_qq} shows that the post filter test statistics looked reasonably Normal. Because the empirical null mean and empirical null standard deviation are random variables with some distribution, the post filtered test statistics will never be Normal and any Normality tests would be too strict for the purpose of this experiment.

\begin{figure}[htp]
	\centering
	\includegraphics[width=0.49\textwidth]{../figures/inference/noContamination/allNullGaussianqq.eps}
	\caption{QQ plot of the test statistics of a filtered $\addNumber{../figures/inference/noContamination/imageSize.txt} \times \addNumber{../figures/inference/noContamination/imageSize.txt}$ Gaussian image after filtering with kernel radius \addNumber{../figures/inference/noContamination/radius.txt}.}
	\label{fig:inference_noContamination_qq}
\end{figure}

\begin{figure}[pht]
	\centering
	\centerline{
	\begin{subfigure}[b]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../figures/inference/noContamination/allNullGaussianunfiltered.eps}
		\caption{Unfiltered}
	\end{subfigure}
	\begin{subfigure}[b]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../figures/inference/noContamination/allNullGaussianfiltered.eps}
		\caption{Filtered}
	\end{subfigure}
	}
	\centerline{
	\begin{subfigure}[b]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../figures/inference/noContamination/allNullGaussiannullmean.eps}
		\caption{Empirical null mean}
	\end{subfigure}
	\begin{subfigure}[b]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../figures/inference/noContamination/allNullGaussiannullstd.eps}
		\caption{Empirical null std}
	\end{subfigure}
	}
	\caption{A $\addNumber{../figures/inference/noContamination/imageSize.txt} \times \addNumber{../figures/inference/noContamination/imageSize.txt}$ Gaussian image before and after filtering with kernel radius \addNumber{../figures/inference/noContamination/radius.txt}. Also shown are the empirical null mean and empirical null std.}
	\label{fig:inference_noContamination_images}
\end{figure}

@@

\begin{figure}[htp]
	\centering
	\begin{subfigure}[b]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../figures/inference/allnull/Experiment_AllNullGaussianmean.eps}
		\caption{Post filter sample mean}
	\end{subfigure}
	\begin{subfigure}[b]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../figures/inference/allnull/Experiment_AllNullGaussianvariance.eps}
		\caption{Post filter sample variance}
	\end{subfigure}
	\caption{The sample mean and sample variance of the test statistics in a filtered Gaussian image of size $\addNumber{../figures/inference/allnull/Experiment_AllNullGaussianwidth.txt} \times \addNumber{../figures/inference/allnull/Experiment_AllNullGaussianheight.txt}$. The box plots summarise the \addNumber{../figures/inference/allnull/Experiment_AllNullGaussiannrepeat.txt} repeats of the experiment. The dotted lines show the critical region at the 2$\sigma$ level using standard tests for the sample mean and sample variance, assuming independence.}
	\label{fig:inference_allnullgaussian}
\end{figure}

The results are shown in Figure \ref{fig:inference_allnullgaussian}. The post filter sample test statistic means do agree with zero as expected. There is some bias with the post filter sample test statistics variances, it was found that this was sensitive to the bandwidth parameters for the kernel density estimate. The large sampling error can come from a number of sources. The empirical null parameters are random variables, thus correction using these can increase the variance. Also pixel to pixel correlation can be introduced as the empirical null parameters are functions of neighbouring pixels, this can contribute to the sample error.

It should be noted that the slight decrease of the post filter test statistic will decrease the sensitivity of the hypothesis test ever so slightly.

%\afterpage{\clearpage}
\subsection{Contaminated Experiment}

Consider a $\addNumber{../figures/inference/contamination/imageSize.txt}\times \addNumber{../figures/inference/contamination/imageSize.txt}$ Gaussian image with some contamination. For example let the pixels be $Z_{x,y}\sim\normal(\mu_{0,x,y},\sigma_0^2)$ where the null mean is a linear function
\begin{equation}
\mu_{0,x,y} = \addNumber{../figures/inference/contamination/gradx.txt} (x-x_0) + \addNumber{../figures/inference/contamination/grady.txt} (y-y_0)
\end{equation}
where $(x_0,y_0)$ is the centre of the image and $\sigma_0=\addNumber{../figures/inference/contamination/nullstd.txt}$. There are no defects or non-null pixels defined here, therefore any positive detections are false.

@@

The QQ plot of the filtered test statistics, as shown in Figure \ref{fig:inference_contamination_qq}, shows that there may be a variance mismatch. A mismatch in the variance will increase or decrease the sensitivity of the test, a larger variance would increase the sensitivity.

\begin{figure}[htp]
	\centering
	\includegraphics[width=0.49\textwidth]{../figures/inference/contamination/allNullPlaneqq.eps}
	\caption{QQ plot of the test statistics of a filtered $\addNumber{../figures/inference/noContamination/imageSize.txt} \times \addNumber{../figures/inference/noContamination/imageSize.txt}$ contaminated Gaussian image after filtering with kernel radius \addNumber{../figures/inference/noContamination/radius.txt}.}
	\label{fig:inference_contamination_qq}
\end{figure}

\begin{figure}[pht]
	\centering
		\centerline{
			\begin{subfigure}[b]{0.49\textwidth}
				\includegraphics[width=\textwidth]{../figures/inference/contamination/allNullPlaneunfiltered.eps}
				\caption{Unfiltered}
			\end{subfigure}
			\begin{subfigure}[b]{0.49\textwidth}
				\includegraphics[width=\textwidth]{../figures/inference/contamination/allNullPlanefiltered.eps}
				\caption{Filtered}
			\end{subfigure}
		}
		\centerline{
			\begin{subfigure}{0.49\textwidth}
				\includegraphics[width=\textwidth]{../figures/inference/contamination/allNullPlanenullmean.eps}
				\caption{Empirical null mean}
			\end{subfigure}
			\begin{subfigure}{0.49\textwidth}
				\includegraphics[width=\textwidth]{../figures/inference/contamination/allNullPlanenullstd.eps}
				\caption{Empirical null std}
			\end{subfigure}
		}
	\caption{A $\addNumber{../figures/inference/contamination/imageSize.txt} \times \addNumber{../figures/inference/contamination/imageSize.txt}$ contaminated Gaussian image before and after filtering with kernel radius \addNumber{../figures/inference/contamination/radius.txt}. The contamination is such that the null distribution is $Z_{x,y}\sim\normal(\mu_{0,x,y},\sigma_0^2)$ where $\mu_{0,x,y} = \addNumber{../figures/inference/contamination/gradx.txt} (x-x_0) + \addNumber{../figures/inference/contamination/grady.txt} (y-y_0)$ and $\sigma_0=\addNumber{../figures/inference/contamination/nullstd.txt}$. Highlighted in red are significant pixels when controlling the FDR at the $2\sigma$ level. Also shown are the empirical null mean and empirical null std.}
	\label{fig:inference_contamination_images}
\end{figure}

@@

The results are shown in Figure \ref{fig:inference_allnullplane}. The filter correctly centred the test statistics at zero. However there was some variance mismatch and this depended on the kernel radius. A kernel radius of about 80 scaled the test statistics correctly. This suggest there may be an optimal kernel radius for a null distribution. In addition, the sensitivity may not be exactly preserved after filtering.

\begin{figure}[htp]
	\centering
	\begin{subfigure}[b]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../figures/inference/allnull/Experiment_AllNullPlanemean.eps}
		\caption{Post filter sample mean}
	\end{subfigure}
	\begin{subfigure}[b]{0.49\textwidth}
		\includegraphics[width=\textwidth]{../figures/inference/allnull/Experiment_AllNullPlanevariance.eps}
		\caption{Post filter sample variance}
	\end{subfigure}
	\caption{The sample mean and sample variance of the test statistics in a filtered contaminated Gaussian image of size $\addNumber{../figures/inference/allnull/Experiment_AllNullGaussianwidth.txt} \times \addNumber{../figures/inference/allnull/Experiment_AllNullGaussianheight.txt}$. The contamination is such that the null distribution is $Z_{x,y}\sim\normal(\mu_{0,x,y},\sigma_0^2)$ where $\mu_{0,x,y} = \addNumber{../figures/inference/contamination/gradx.txt} (x-x_0) + \addNumber{../figures/inference/contamination/grady.txt} (y-y_0)$ and $\sigma_0=\addNumber{../figures/inference/contamination/nullstd.txt}$. The box plots summarise the \addNumber{../figures/inference/allnull/Experiment_AllNullPlanenrepeat.txt} repeats of the experiment. The dotted lines show the critical region at the 2$\sigma$ level using standard tests for the sample mean and sample variance, assuming independence.}
	\label{fig:inference_allnullplane}
\end{figure}
